Geometrization of characters of smooth group schemes over finite fields with an application to types for supercuspidal representations

### 2016-02-23 followup.tex
Changes to the introduction.

\title[Commutative Character Sheaves]{Commutative Character Sheaves on smooth group schemes over finite fields}
%with an application to geometric types for supercuspidal representations
\author{Clifton Cunningham}
\address{Department of Mathematics and Statistics, University of Calgary, 2500 University Drive Northwest, Calgary, Alberta, Canada, {T2N~1N4}.}
\email{cunning@math.ucalgary.ca}
\author{David Roe}
\address{Department of Mathematics, University of Pittsburgh, 301 Thackeray Hall, Pittsburgh, PA , United States, 15260.}
\email{roed.math@gmail.com}

@@

\begin{abstract}
We describe a function-sheaf dictionary for one-dimensional characters of arbitrary smooth group schemes
over finite fields.  In previous work, we considered the case of commutative smooth group schemes and found
that the standard definition of character sheaves produced a dictionary with a nontrivial kernel.  In this paper,
we give a modification of the category of character sheaves that remedies this defect, and is also extensible to
non-commutative groups.  
As an application, we use these \emph{commutative character sheaves} to geometrize types for
supercuspidal representations of $p$-adic groups. 
\end{abstract}

@@

Finally, we use the category $\CCS(G)$ to give a geometric description of generic types for supercuspidal representations
of connected reductive groups over local fields.  In this way, we provide all of the ingredients
needed to parametrize supercuspidal representations of arbitrary depth in the same category:
perverse sheaves on group schemes over finite fields. \todo{Update based on any changes in Section 4}

In Section \ref{sec:defs}, we recall the category $\CS(G)$ from \cite{cunningham-roe:13a} and note that it still makes
sense when $G$ is not commutative.  We focus on the case of commutative $G$ in Section \ref{sec:comcom},
giving the definition of a commutative character sheaf and proving our first main theorem, that
$\TrFrob{G} : \CSiso{G} \to G(k)^*$ induces an isomorphism on $\CCSiso{G}$.

Passing to the case that $G$ is
non-commutative, we give the definition of and main results about commutative character sheaves in
Section \ref{sec:noncom}.  We note that we should only consider character sheaves
that arise via pullback from $G\ab$ in order to eliminate those that have nontrivial restriction
to the derived subgroup.  This observation underlies the definition of commutative character sheaves for
non-commutative $G$.  We state our second main theorem, that pullback along the abelianization map
defines an equivalence of categories $\CCS(G) \to \CCS(G\ab)$.
In Section \ref{ssec:obmor}, we use Galois cohomology to describe the relationship between $G(k)^*$
and $G\ab(k)^*$.  We also compute the automorphism groups in $\CCS(G)$.
Then in Section \ref{ssec:proofs}, we give proofs of the results in Section \ref{sec:noncom}, which
require a development of equivariant linear character sheaves.

\todo{Add summary of Section 4 once it has stabilized}

In the remainder of this section we use Theorem~\ref{thm:geo} to geometrize the representation $\rho$ of $C$: subject to a hypothesis on the component group of the depth-zero part of the datum and also on a conjecture on the geometrization of the character of Heisenberg-Weil representations, we find a group scheme $G$ over $\Fq$ such that $G(\Fq) = C$ and a Weil sheaf $\mathcal{F}$ on $G$ such that $\trFrob{\cs{F}} = \trace(\rho)$.
The proof of this statement is quite technical, so it must be delayed to Section~\ref{ssec:geotypes}, after reviewing types in Section~\ref{ssec:types} and models Section~\ref{ssec:models} developed by Yu, for which we must also review some facts about the Heisenberg-Weil representation in Section~\ref{ssec:Jacobi}, and also make a conjecture about the geometrization of that representation, Conjecture~\ref{conj}.

>>>>

\title[Geometrization of characters of smooth group schemes over finite fields]{Geometrization of characters of smooth group schemes over finite fields with an application to types for supercuspidal representations}
\author{Clifton Cunningham}
\address{Department of Mathematics and Statistics, University of Calgary, 2500 University Drive Northwest, Calgary, Alberta, Canada, {T2N~1N4}.}
\email{cunning@math.ucalgary.ca}
\author{David Roe}
\address{Department of Mathematics, University of Pittsburgh, 301 Thackeray Hall, Pittsburgh, PA , United States, 15260.}
\email{roed.math@gmail.com}

@@

\begin{abstract}
We describe a function-sheaf dictionary for one-dimensional characters of arbitrary smooth group schemes over finite fields.  
In previous work, we considered the case of commutative smooth group schemes and found that the standard definition of character sheaves produced a dictionary with a nontrivial kernel.  
In this paper we give a modification of the category of character sheaves that remedies this defect, and is also extensible to non-commutative groups.  
As an application, we use these \emph{commutative character sheaves} to geometrize types for supercuspidal representations of $p$-adic groups, assuming the existence of the geometrization of the Heisenberg-Weil representation over finite fields.
\end{abstract}

@@

Finally, we use the category $\CCS(G)$ to give a geometric description of types for supercuspidal representations of connected reductive groups over $p$-adic fields fields,
assuming the existence of a geomtrization of the Heisenberg-Weil representations over finite fields.
In this way, we provide all of the ingredients needed to parametrize supercuspidal representations of arbitrary depth in the same category:
virtual Weil perverse sheaves on smooth group schemes over finite fields. 
%\todo{Update based on any changes in Section 4}

We now summarize the sections of the paper in more detail.
In Section \ref{sec:defs}, we recall the category $\CS(G)$ from \cite{cunningham-roe:13a} and note that it still makes
sense when $G$ is not commutative.  We focus on the case of commutative $G$ in Section \ref{sec:comcom},
giving the definition of a commutative character sheaf and proving our first main theorem, that
$\TrFrob{G} : \CSiso{G} \to G(k)^*$ induces an isomorphism on $\CCSiso{G}$.
%
Passing to the case that $G$ is non-commutative, we give the definition of and main results about commutative character sheaves in Section \ref{sec:noncom}.  
We note that we should only consider character sheaves that arise via pullback from $G\ab$ in order to eliminate those that have nontrivial restriction to the derived subgroup.  
This observation underlies the definition of commutative character sheaves for non-commutative $G$.  
We state our main result, Theorem~\ref{thm:geo}, that pullback along the abelianization map defines an equivalence of categories $\CCS(G) \to \CCS(G\ab)$.
In Section \ref{ssec:obmor}, we use Galois cohomology to describe the relationship between $G(k)^*$ and $G\ab(k)^*$.  
We also compute the automorphism groups in $\CCS(G)$.
Then in Section \ref{ssec:proofs}, we give proofs of the results in Section \ref{sec:noncom}, which require a development of equivariant linear character sheaves.

In Section~\ref{sec:types} we use Theorem~\ref{thm:geo} to geometrize the representations of compact $p$-adic groups appearing in the types for : subject to a hypothesis on the component group of the depth-zero part of the datum and also on a conjecture on the geometrization of the character of Heisenberg-Weil representations, we find a group scheme $G$ over $\Fq$ such that $G(\Fq) = C$ and a Weil sheaf $\mathcal{F}$ on $G$ such that $\trFrob{\cs{F}} = \trace(\rho)$.
The proof of this statement is quite technical, so it must be delayed to Section~\ref{ssec:geotypes}, after reviewing types in Section~\ref{ssec:types} and models Section~\ref{ssec:models} developed by Yu, for which we must also review some facts about the Heisenberg-Weil representation in Section~\ref{ssec:Jacobi}, and also make a conjecture about the geometrization of that representation, Conjecture~\ref{conj}.
